---
title: "Exercise #3"
subtitle: "Fortgeschrittene Statistische Software für NF"
author: "Emily Burton (12597288)"
date: "`r Sys.Date()`"
output: distill::distill_article
---

## General Remarks

-   You can submit your solutions in teams of up to 3 students.
-   Include all your team-member's names and student numbers
    (Matrikelnummern) in the `authors` field.
-   Please use the exercise template document to work on and submit your
    results.
-   Use a level 2 heading for each new exercise and answer each subtask
    next to its bullet points or use a new level 3 heading if you want.
-   Always render the R code for your solutions (`echo=TRUE`) and make
    sure to include the resulting data in your rendered document.
    -   Make sure to not print more than 10 rows of data (unless
        specifically instructed to).
-   Always submit both the rendered document(s) as well as your source
    Rmarkdown or Quarto document. Submit the files separately on moodle,
    **not** as a zip archive.
-   Submission format is HTML. Other formats will lead to a deduction of
    points.

## Exercise 1: Initializing git (4 Points)

For this whole exercise sheet we will be tracking all our changes to it
in git.

a)  Start by initializing a new R project with git support, called
    `2025-exeRcise-sheet-3`. If you forgot how to do this, you can
    follow this
    [guide](https://malikaihle.github.io/Introduction-RStudio-Git-GitHub/rstudio_project.html).
b)  Commit the files generated by Rstudio.
c)  For all of the following tasks in this exercise sheet we ask you to
    always commit your changes after finishing each subtask e.g. create
    a commit after task *1d*, *1e* etc.

> Note: This applies only to answers that have text or code as their
> answer. If you complete tasks in a different order or forget to commit
> one, this is no problem. If you change your answers you can just
> create multiple commits to track the changes.

d)  Name 2 strengths and 2 weaknesses of git. (Don't forget to create a
    commit after this answer, see *1c*)
    
    ### 1d) Strengths and Weaknesses of Git

**Strengths:**
1. Version control: this allows you to track changes and revert to previous states easily.
2. It supports collaboration, enabling multiple people to work on the same project without overwriting each other's work.

**Weaknesses:**
1. Git has a steep learning curve, especially for users unfamiliar with command-line tools or version control systems.
2. Merge conflicts can be confusing and difficult to resolve, especially in larger projects or for beginners.

e)  Knit this exercise sheet. Some new files will automatically be
    generated when knitting the sheet e.g. the HTML page. Ignore these
    files, as we only want to track the source files themselves. You
    can, but don't need to create a `.gitignore` file. Just do not
    commit these files manually.

## Exercise 2: Putting your Repository on GitHub (3 Points)

For this task you will upload your solution to GitHub.

a)  Create a new repository on GitHub in your account named
    `exeRcise-sheet-3`. Make sure you create a **public repository** so
    we are able to see it for grading. Add the link to the repository
    below:
b)  Push your code to this new repository by copying and executing the
    snippet on github listed under
    `…or push an existing repository from the command line`.
c)  Regularly push your latest changes to GitHub again and especially do
    so when you are finished with this sheet.

## Exercise 3: Pixar Films (4 Points)

Download the `pixar_films` and `public_response` datasets from the
GitHub repository and track them in git.

Link:
<https://github.com/rfordatascience/tidytuesday/tree/main/data/2025/2025-03-11>

For small datasets like these adding them to git is not a problem.

a)  Load the `pixar_films` dataset into R. Clean the dataset by removing
    films without a title. Inspect the variable `film_rating`. What are
    the possible values and what do they mean? Create a factor variable
    for the film rating. Why is this appropriate?
    
```{r}
library(tidyverse)
library(readr)
library(dplyr)

# Load data directly from GitHub
pixar_films <- read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/main/data/2025/2025-03-11/pixar_films.csv')
public_response <- read_csv('https://raw.githubusercontent.com/rforddatascience/tidytuesday/main/data/2025/2025-03-11/public_response.csv')

# Clean pixar_films: remove films without a title
pixar_films <- pixar_films %>% filter(!is.na(film))

# film_rating variable: what are the possible values?
ratings <- unique(pixar_films$film_rating)
print(ratings)

# Explanation of film_rating values:
# "G"   - General audiences, suitable for all ages.
# "PG"  - Parental guidance suggested.
# "N/A" - Not available or not rated.

# Convert film_rating to a factor
pixar_films <- pixar_films %>%
  mutate(film_rating = factor(film_rating))

# Check first rows
head(pixar_films)

# The film_rating variable represents categories (e.g., “G”, “PG”, “N/A”) that classify films into groups rather than numeric values. Factors are the appropriate data type in R for categorical variables because they allow statistical functions and plots to correctly treat these values as categories.

```



b)  Inspect the film titles manually. Which films form a film series? A
    film series can be identified by a common word in the titles of the
    films, often in conjunction with a number in the title,
    e.g. "Despicable Me" and "Despicable Me 2". Create a dataframe which
    displays a list of the different series with the titles of the films
    and how many films belong to the series. Output the dataframe.
    
```{r}
library(tidyverse)
library(stringr)


colnames(pixar_films)
print(pixar_films$film)

pixar_films <- pixar_films %>%
  mutate(series = case_when(
    str_detect(film, "Toy Story") ~ "Toy Story",
    str_detect(film, "Cars") ~ "Cars",
    str_detect(film, "Finding") ~ "Finding",
    str_detect(film, "Incredibles") ~ "Incredibles",
    TRUE ~ NA_character_
  ))

# Summary 
film_series_df <- pixar_films %>%
  filter(!is.na(series)) %>%
  group_by(series) %>%
  summarise(
    films = paste(film, collapse = ", "),
    num_films = n(),
    .groups = "drop"
  )

print(film_series_df)
```
c)  Load the `public_response` dataframe into R. Convert the
    `cinema_score` variable into a factor while ensuring the factor
    levels are defined in ascending order, from the lowest to the
    highest score. Combine `public_response` with the `pixar_films`
    dataset using an appropriate merge variable.
    
```{r}

# Check columns
print(colnames(pixar_films))
print(colnames(public_response))

# Convert cinema_score into ordered factor
cinema_levels <- c("F", "D-", "D", "D+", "C-", "C", "C+", "B-", "B", "B+", "A-", "A", "A+")
public_response <- public_response %>%
  mutate(cinema_score = factor(cinema_score, levels = cinema_levels, ordered = TRUE))

# Select only necessary columns from public_response to avoid duplicates
public_response_sel <- public_response %>%
  select(film, cinema_score, rotten_tomatoes, metacritic, critics_choice)

# Merge datasets by title (pixar_films) and film (public_response)
combined_df <- left_join(pixar_films, public_response_sel, by = c("title" = "film"))

# Check which films from pixar_films are missing in public_response
missing_films <- pixar_films$title[!pixar_films$title %in% public_response_sel$film]
print(missing_films)  # Films without matching public_response data

# View result
head(combined_df)
```

d)  Choose one of the variables representing the public response and
    create a bar plot for the films belonging to a series. Here are the
    details of the plot:

    -   The film series are represented on the x-axis.
    -   Your chosen public response variable is displayed on the y-axis.
    -   Each film in the series is represented as a separate bar. Bars
        are grouped by film under their respective series on the x-axis.
        Order the bars within a series according to the release date of
        the films.
    -   A title and axis labels for context.

    What do you notice when comparing the scores of the films in a
    series? Do you see any patterns?
```{r}
# Filter films that are part of a series and have cinema_score data
plot_data <- combined_df %>%
  filter(!is.na(series), !is.na(cinema_score)) %>%
  arrange(series, release_date) %>%
  mutate(title = factor(title, levels = unique(title)))

# Plot
ggplot(plot_data, aes(x = series, fill = title, y = as.numeric(cinema_score))) +
  geom_bar(stat = "identity", position = position_dodge(width = 0.9)) +
  scale_y_continuous(breaks = 1:length(levels(public_response$cinema_score)),
                     labels = levels(public_response$cinema_score)) +
  labs(
    title = "Cinema Scores by Film Series",
    x = "Film Series",
    y = "Cinema Score (Ordered)",
    fill = "Film Title"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

The cinema scores show clear patterns across film series.
Sequels in the Cars and Finding series tend to receive slightly lower scores than the original films.
However, across all series, Pixar films generally maintain strong audience approval.

## Exercise 4: Open Analysis (4 points)

This exercise is a bit more open-ended. You can choose any dataset from
[Our World in Data](https://ourworldindata.org/) and analyze it, while
determining the research question yourself.



a)  Go to <https://github.com/owid/owid-datasets/tree/master/datasets>
    and choose a dataset that interests you. You can have a look at
    <https://ourworldindata.org/> to gather some inspiration.
    
Dataset chosen: https://github.com/owid/owid-datasets/tree/master/datasets/CO2%20Footprint%20Breakdown%20Per%20Capita%20-%20Goodall%20(2011)

```{r}
# Load tidyverse
library(tidyverse)


co2_lifestyle <- read.csv("/Users/emilyburton/2025-exeRcise-sheet-3/co2_footprint.csv")


head(co2_lifestyle)
```
b)  Download the dataset and track it in git.
c)  Put the name / title of the dataset and a link to it below.

-   Dataset Name: CO2 Footprint Breakdown Per Capita (Goodall 2011)
-   Link: <https://github.com/owid/owid-datasets/blob/master/datasets/CO2%20Footprint%20Breakdown%20Per%20Capita%20-%20Goodall%20(2011)/CO2%20Footprint%20Breakdown%20Per%20Capita%20-%20Goodall%20(2011).csv>

d)  Come up with a (research) question you want to answer with the data
    and briefly explain why you believe this is an interesting question
    within one sentence. It should be a question that can be answered
    with the dataset and using R.
    
    Question:
    Which sector contributes the most to per capita CO2 footprint in 2011, 
    and how do the contributions compare across sectors?

    Why:
    Understanding sector contributions helps target policies to reduce emissions effectively.
    
e)  Use R to answer your chosen question. Please limit your analysis to
    the functions and techniques we have covered so far in the course.
    You are **not expected** to use advanced statistical models or
    external packages which haven't been introduced.
    
    ```{r}
df <- read.csv("co2_footprint.csv", stringsAsFactors = FALSE)

# View the structure
str(df)

names(df) <- c("Entity", "Year", "Percent")
df$CO2.Percent <- as.numeric(gsub("%", "", df$Percent))

# Find the sector with the highest CO2 contribution
max_sector <- df[which.max(df$CO2.Percent), ]
print(max_sector)
    ```
f)  Create a meaningful plot / figure with the dataset. Make sure to
    provide a figure caption (via the chunk options / Rmarkdown) and
    correctly label the figure.
```{r}
barplot(df$CO2.Percent,
        names.arg = df$Entity,
        las = 2,                      # Rotate labels
        col = "skyblue",
        main = "CO2 Footprint Breakdown per Capita by Sector (2011)",
        ylab = "CO2 Contribution (%)",
        cex.names = 0.8)

text(x = seq_along(df$CO2.Percent),
     y = df$CO2.Percent,
     labels = paste0(df$CO2.Percent, "%"),
     pos = 3, cex = 0.75, col = "darkblue")
```
## Final Note

Make sure to push all your commits and changes to GitHub before
submitting the exercise sheet.
